---
output: html_document
---

###Load required packages
```{r, message = FALSE, warning = FALSE}
library(data.table) #reading in the data
library(dplyr) #dataframe manipulation
library(ggplot2) #viz
library(ranger) #the random forest implementation
library(plotly) #3D plotting
library(tidyr) #dataframe manipulation
library(FNN) #k nearest neighbors algorithm
library(xgboost)
```

###Load data, count total #, view columns
```{r}
fb <- fread("~/PycharmProjects/kaggle-project/facebook/train.csv", integer64 = "character", showProgress = FALSE)
nrow(fb)
```

```{r}
head(fb, 3)
```

```{r}
summary(fb)
```


###Abstract sample area
```{r}
fb %>% filter(x >1, x <1.25, y >2.5, y < 2.75) -> fb_s
nrow(fb_s)  
```

```{r}
fb %>% filter(x >3, x <3.25, y >2.5, y < 2.75) -> fb_s2
nrow(fb_s2)  
```

##Observe data
###place_id
**Since target is to classify place_id, 1st to observe place_id**
From figures: horizon vs continuously increase
```{r}
par(mfrow=c(3,1))
plot(sort(fb_s$place_id))
plot(sort(fb_s2$place_id))
plot(sort(fb_s2$place_id)[0:2000])
```

From figures: seem no relation between place_id and accuracy
```{r}
par(mfrow=c(2,1))
r_pla_aur=sort(fb_s2$place_id,index.return=TRUE)
plot(r_pla_aur$x[0:2000])
d=r_pla_aur$ix[0:2000]
plot(fb_s2[d,"accuracy"])
```

```{r}
par(mfrow=c(2,1))
r_pla_aur=sort(fb_s2$place_id,index.return=TRUE)
plot(r_pla_aur$x[500:750])
d=r_pla_aur$ix[500:750]
plot(fb_s2[d,"accuracy"])
```


###time
```{r}
fb_s$hour = (fb_s$time/60) %% 24
fb_s$weekday = (fb_s$time/(60*24)) %% 7
fb_s$month = (fb_s$time/(60*24*30)) %% 12 #month-ish
fb_s$year = fb_s$time/(60*24*365)
fb_s$day = fb_s$time/(60*24) %% 365
head(fb_s)
```

##Train model
**Split data**
```{r}
small_train = fb_s[fb_s$time < 7.3e5,]
small_val = fb_s[fb_s$time >= 7.3e5,] 
```

**visualize 2D: small_train**
```{r, fig.height = 8, fig.width = 10}
ggplot(small_train, aes(x, y )) +
    geom_point(aes(color = place_id)) + 
    theme_minimal() +
    theme(legend.position = "none") +
    ggtitle("Check-ins colored by place_id")
```

Count by place_id
```{r}
sort((small_train %>% count(place_id))$n, decreasing = T)[0:140]
```

**visualize 3D: small_train with place_id count>500**
**z=hour**
```{r, fig.height = 8, fig.width = 8}
small_train %>% count(place_id) %>% filter(n > 500) -> ids
  #if n>200, warning: n too large, allowed maximum for palette Set2 is 8
small_trainz = small_train[small_train$place_id %in% ids$place_id,]

plot_ly(data = small_trainz, x = x , y = y, z = hour, color = place_id,  type = "scatter3d", mode = "markers", marker=list(size= 5)) %>% layout(title = "Place_id's by position and Time of Day")
```

**z=week**
```{r}
plot_ly(data = small_trainz, x = x , y = y, z = weekday, color = place_id,  type = "scatter3d", mode = "markers", marker=list(size= 5)) %>% layout(title = "Place_id's by position and Day of Week")

```


##Random Forest

**Count unique place_id**
```{r}
length(unique(small_train$place_id))
```

**Ignore fewer place_id**
```{r}
small_train %>% count(place_id) %>% filter(n > 3) -> ids
small_train = small_train[small_train$place_id %in% ids$place_id,]
```

#optimal weights for scaling your variables since knn is sensitive to the magnitutde of variables@¥: s,l,w
```{r}
summary(small_train)
```

```{r}
s = 2
l = 125
w = 500

create_matrix = function(train) {
    cbind(s*train$y,
          train$x,
          train$hour/l,
          train$weekday/w,
          train$year/w,
          train$month/w,
          train$time/(w*60*24*7))
    }

X = create_matrix(small_train)
X_val = create_matrix(small_val)

```

**KNN**
```{r}
model_knn = FNN::knn(train = X, test = X_val, cl = small_train$place_id, k = 15)

preds <- as.character(model_knn)
truth <- as.character(small_val$place_id)
mean(truth == preds)
```

```{r}
head(X)
```


**Random Forest**
```{r}
set.seed(131L)
small_train$place_id <- as.factor(small_train$place_id) # ranger needs factors for classification
model_rf <- ranger(place_id ~ x + y + accuracy + hour + weekday + month + year,
                   small_train,
                   num.trees = 100,
                   write.forest = TRUE,
                   importance = "impurity")


pred = predict(model_rf, small_val)
pred = pred$predictions
accuracy = mean(pred == small_val$place_id) 

accuracy
```

**Visualize RF accuracy**
It does seem that the correctly identified check-ins are more "clustered" while the wrongly identified ones are more uniformly distributed but other than that no clear patters here.
```{r}
small_val$Correct = (pred == small_val$place_id)

ggplot(small_val, aes(x, y )) +
    geom_point(aes(color = Correct)) + 
    theme_minimal() +
    scale_color_brewer(palette = "Set1")
```

**look at what kind of id's our random forest gets wrong**
We see below that our model is doing actually really great on the more popular id's(more blue on the right). However it loses when it looks at id's that appear only a few times. 
```{r, fig.width = 12}
#reordering the levels based on counts:
small_val$place_id <- factor(small_val$place_id,
                             levels = names(sort(table(small_val$place_id), decreasing = TRUE)))

small_val %>% 
    ggplot(aes(x = place_id)) + geom_bar(aes(fill = Correct)) + 
    theme_minimal() +
    theme(axis.text.x = element_blank()) +
    ggtitle("Prediction Accuracy by ID and Popularity") +
    scale_fill_brewer(palette = "Set1")
```

**importance of our variables**
1. `y` variable is more important than the `x`
This means that the `y` axis is a better predictior of `place_id` and the random forest figures this out on its own. 
2. `hour` and other time features are also good predictiors but less so than the spatial features - this makes sense since the location of a check-in should be more important than the time of the check-in.
3. Accuracy is a bit misterious since we don't get an actual definition for it, but at least the model tells us it's somewhat important.
```{r}
data.frame(as.list(model_rf$variable.importance)) %>% gather() %>% 
    ggplot(aes(x = reorder(key, value), y = value)) +
    geom_bar(stat = "identity", width = 0.6, fill = "grey") +
    coord_flip() +
    theme_minimal() +
    ggtitle("Variable Importance (Gini Index)") +
    theme(axis.title.y = element_blank()) 

```





